Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-01-08 Manuscript/TEM_draft.tex

% NC: Quick comment. You find that Bayesian consensus trees are "best". BUT I think you need to be careful here with what you mean by "best". They are most likely to recover a more similar topology to the best tree than if you do the crazy tree comparison pairs. But that doesn't mean they're best, just that they perform best in your simulations. They are only best for that. If you want to advise people to use a method that best recovers topology as measured using RF or Triplets, then Bayesian consensus is best. BUT generally that is not what people want. Probably need to think some more about this.
% TG: a note about the "best" tree topology regarding your comment just after the discussion header: should I replace "correct" by "best" since that's how we're calling the "best" tree anyway (all in quotes)

@@

Our results show that the ability to recover the "best" tree topology in a Total Evidence framework decreases with increasing amount of missing data, regardless of how data were removed or the method of tree inference used. However, these factors affected topological recovery in different ways and to different extents. 
Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}). Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).As proposed in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough morphological characters overlaps for most living and fossil taxa in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}. % NC: What do you mean by data overlap here? That species have overlapping characters?

\subsection{Individual effects of missing data parameters}
\subsubsection{Effect of missing data for living taxa}
When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix. Because living taxa still have 10 times more molecular characters available (see methods) %link to methods
, even if they have no morphological data, the relation between them will always be fairly well resolved (depending on the phylogenetic signal from the molecular part of the matrix) and thus, ignoring the fossils, the living taxa clades will be always well conserved. However, this missing data parameter influences a lot the placement of fossil taxa because a decrease of the $M_{L}$ parameter will reduce the overlapping data between the living and the fossil taxa.

When using the Normalised Robinson-Foulds distance to measure topological similarity, the Bayesian consensus trees recover the "best" tree topology more often than Maximum Likelihood trees (Fig.~\ref{Fig_Results-global_perparam}). In fact, regardless of the amount of missing data, Bayesian consensus trees always displays the highest Normalised Robinson-Foulds score (always $>$ 0.69; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for all the values). However, regarding the Normalised Triplets distance, the similarity with the "best" tree topology decreases in the same way for both the Maximum Likelihood trees and the Bayesian consensus trees for this particular missing data parameter (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
% TG: Or is that line just repeating the results?
However, regardless of the metric, the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution always performed poorlier in their ability to recover the "best" tree topology, even when no morphological data was missing ($M_{L}$ = 0\% - Fig.~\ref{Fig_Results-permeth_perparam}). This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" trees and a set of the "best" trees involved 1000 random pairwise comparisons rather than just one. This will obviously add noise to the results for this method of measuring the ability to recover the "best" tree topology.

Variation in our ability to recover the "best" tree topology depends heavily on the tree inference method (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}). However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} for recovering the "best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework. Our results shows that the topology of the Bayesian consensus tree is always closer to the "best" tree topology than the one of the Maximum Likelihood tree (Fig.~\ref{Fig_Results-global_perparam}).

In the Bayesian consensus tree topology, the clade conservation (i.e. approximated by the Normalised Robinson-Foulds distance) stays always high but the placement of wildcard taxa (i.e. approximated by the Normalised Triplets distance) decreases way faster (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). The difference between the metrics is linked to the way each is calculated and to how they perform at underlining changes in topology (see method %add link
, \hyperref[SupplementaryMaterial]{Supplementary Material S2} and \citealt{kuhnerpractical2014} for details on the nature of both metrics). In general, the Normalised Robinson-Foulds distance is more sensitive than the Normalised Triplets distance.
%----> version one
When getting closer to the root of the tree, the displacement of a single taxon will have a greater effect than it would in higher clades and therefore decrease the Normalised Robinson-Foulds distance.
%----> version two
%When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
Also, \citet{kuhnerpractical2014} have demonstrated that the Robinson-Foulds distance outperforms the Triplets distance for measuring topological differences between trees. Thus, the Bayesian consensus tree topology seems to be the less affected by removing morphological data for living taxa (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).

\subsubsection{Effect of missing data for fossil taxa}
When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this also reduces the probability of morphological characters for the fossil taxa to overlap with the ones for the living taxa. This can lead to difficulties for certain taxa to be branched in the tree (i.e. wildcard taxa behaviour). However, it is important to note that even though the number of displaced wildcard taxa increases (i.e. decrease of Normalised Triplets distance), the conservation of the clades (i.e. Normalised Robinson-Foulds distance) is still relatively good (mode = @% value
) when the proportion of missing data is high ($M_{F}$ = 75\%).

%----> version one
The effect of the missing data in the fossil record ($M_{F}$) on the Normalised Robinson-Foulds distance (i.e. as a proxy for clade conservation) is smaller than the one of the $M_{L}$ parameter in the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). However, the effect of the $M_{F}$ parameter on the Normalised Triplets distance (i.e. as a proxy for displacement of wildcard taxa) is greater than the one of the $M_{L}$ parameter in the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
%----> version two
%The effect of the missing data in the fossil record ($M_{F}$) is smaller than the one the $M_{L}$ parameter for the clade conservation (Normalised Robinson-Foulds distance) but greater for the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
This can be related to the fact that the Bayesian consensus tree is built using a majority consensus rule. When a fossil taxa have less data (e.g. ($M_{F}$ = 75\%) they will tend to branch "randomly" to any taxa in the clade sharing the most characters. Therefore a majority consensus position is unlikely to exist (i.e. every branching position is represented in $<$ 50\% of the trees) and the fossil taxa will be branched on a polytomy at the base of the clade. This does not decreases the Normalised Robinson-Foulds distance (clades is conserved) but increases the Normalised Triplets distance (wildcard taxa is displaced). On the opposite, the same scenario in a Maximum Likelihood will lead to a dichotomous branching of the fossil but with a low bootstrap support ($<$ 50). In other words, the Bayesian consensus tree allows a fossil taxa with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree (where the fossil will be placed with lower confidence at a higher taxonomic position). We can argue that the first option (the Bayesian consensus tree topology) is preferable \cite[e.g][]{}. %Cite the Darwinius debate or is that a bit dangerous?

\subsubsection{Effect of missing character ($M_{C}$)}
The overall number of missing morphological characters is expected to have an effect on our ability to recover the "best" tree topology two times bigger than for the $M_{L}$ and $M_{F}$ parameters. For example, removing 10\% of the data for the fossil taxa actually removes a total of 5\% of data in the whole matrix (because this parameter affects only half of the taxa present in the matrix); on the other hand, removing 10\% of morphological characters actually removes 10\% of data in the matrix. However the effect of removing characters on the ability to recover the "best" tree topology is of the same order of magnitude than for the two other parameters (Fig.~\ref{Fig_Results-permeth_perparam}). Reducing the overall amount of morphological characters reduces the probability of overlap of morphological characters between the taxa present in the matrix and therefore decreases our ability to recover the "best" tree topology.

%I moved the stuff about simulating characters over here.
The overall number of morphological character in a Total Evidence matrix can also affect our ability to recover the "best" tree topology through two phylogenetic caveats from morphological data (in opposition to molecular data): (1) the incongruence of phylogenetic signal between morphological and molecular data and (2) homoplasy. The incongruence of phylogenetic signal between morphological and molecular data as previously been demonstrated as being potentially more important in small morphological matrices \citep{wagner2000}. Therefore, part of the decrease of the Normalised Robinson-Fould distance and the Normalised Triplets distance in our simulation could be due to conflicting phylogenetic signal between morphological and molecular data (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013,slaterphylogenetic2013,beckancient2014}. (2) Homoplasy, on the other hand is expected to increase with an increase of the morphological characters \citep{wagner2000}. %CITE
However, the use of probabilistic methods (i.e. Maximum Likelihood or Bayesian) and the M\textit{k} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.

\subsection{Combined effect of missing data parameters}
As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the effect of the missing data parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance is higher when all the missing data parameters have low missing data (i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have high missing data (i.e. $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-global_perparam}). However, it is important to notice that the effect of each parameter is not additive. The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological character ($M_{C}$) have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$); when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}).

\subsection{Suggested applications} %Or "Practical implications"
Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections. Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "best" tree topology. We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections. Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries. However, gaps in the matrix can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}. % TG: I added the palaeo field work just as a kind work to the pokemon masters that could read this paper - maybe useless?
Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "best" tree topology (Fig.~\ref{Fig_Results-paircomp_within}). Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved. One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}, so that no one individual collects all the data.

%From here I'm not quite sure how to put that and if it's actually important.
% ---->
Our simulation were constraint by the performance of our protocol: in order to reduce the computational time of our analysis (150 CPU years), we ran our sensitivity analysis on modestly sized matrices of 1000 molecular characters and 100 morphological characters. These matrices are two orders of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013} but still ranges in the sizes of more modest ones \citep[e.g.][]{sallam2011craniodental,kellymolecular2014}. Another point that was constrained by the performance of our protocol was the ability to assess the effect of missing data on branch length variation in a Total Evidence framework which is a crucial parameter in one of the wider applications of matrices combining both living and fossil taxa: the Tip-dating method \citep[e.g.][]{ronquista2012,BEASTmaster}.

Also, simulating morphological characters is more complex than molecular ones. The underlying pattern of their evolution is often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001,wrightbayesian2014}. Morphological characters studies involve many potential statistical pitfalls (e.g. independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy that are discussed above. % add link
Secondly, in this study, we made the assumption that morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013,pattinsonphylogeny2014}. Approaches such as artificial fossilisation \citep{pattinsonphylogeny2014} might solve this issue and 
% <----

@@

Using Total Evidence matrices with missing data is not a problem for recovering the "best" tree topology as long as enough morphological characters are overlapping between the fossil and living taxa. When missing data increases in any of our missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa. Therefore we advise filling as many gaps in Total Evidence matrices as possible. Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters as possible and to code them for a maximum of living taxa present in the matrix. Additionally, the topology of the Bayesian consensus tree, regardless the amount of missing data, was always the closer to the "best" tree topology than the Maximum likelihood trees. Therefore, we advise to use the topology of the Bayesian consensus tree over the topology of the Maximum Likelihood tree or of a random subset of the Bayesian posterior trees distribution

>>>>

@@

Our results show that the ability to recover the "best" tree topology in a Total Evidence framework decreases with increasing amount of missing data, regardless of how data were removed or the method of tree inference used. However, these factors affected topological recovery in different ways and to different extents. 
Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}). Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}). As shown in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.

\subsection{Individual effects of missing data parameters}
\subsubsection{Missing data for living taxa ($M_{L}$)}
When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix. Because living taxa still have molecular characters available for phylogenetic inference (see methods), %link to methods
even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).  However, this missing data parameter has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.

\subsubsection{Missing data for fossil taxa ($M_{F}$)}
When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this also reduces the probability of morphological characters for fossil taxa overlapping with the ones for living taxa. This can lead to difficulties for the placement of certain taxa in the tree. However, it is important to note that even though the number of displaced wildcard taxa increases (i.e. decrease of Normalised Triplets distance) with increasing missing data in this parameter, clade conservation (i.e. Normalised Robinson-Foulds distance) is still relatively good (mode = @) when the proportion of missing data is high ($M_{F}$ = 75\%).

The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds distance) but greater on the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule. When the fossil taxa have less data (e.g. $M_{F}$ = 75\%) they will tend to branch with any taxon in the clade that shares most characters with the fossils. Therefore a majority consensus position is unlikely to exist (i.e. every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade. This does not decrease the Normalised Robinson-Foulds distance, because clades are conserved, but decreases the Normalised Triplets distance because the fossils act as wildcard taxa. Conversely, the same scenario in a Maximum Likelihood framework will lead to a dichotomous branching of the fossils but with low bootstrap support ($<$ 50). In other words, the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree, where the fossil will be placed with lower confidence at a higher taxonomic level. We argue that using the Bayesian consensus tree topology is preferable because it is more conservative \cite[e.g][]{}. %Cite the Darwinius debate or is that a bit dangerous?
% NC: Check pattinson

\subsubsection{Missing morphological characters ($M_{C}$)}
Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the "best" tree topology. We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix). Conversely, removing 10\% of morphological characters ($M_{C}$) genuinely removes 10\% of data in the matrix. However, the effect of removing characters on the ability to recover the "best" tree topology is of the same order of magnitude as for the other two parameters (Fig.~\ref{Fig_Results-permeth_perparam}). We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se} % see above or below???.

Additionally, the number of morphological characters determines the size the matrix. This can affect our ability to recover the "best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy. The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \citep{wagner2000}. The size of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters. Therefore, part of the decrease of the Normalised Robinson-Fould distance and the Normalised Triplets distance in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). However, although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}. Therefore, our simulations reflect realistic parameters. Homoplasy, on the other hand, is expected to increase with an increase of the morphological characters \citep{}. However, the use of probabilistic methods (i.e. Maximum Likelihood or Bayesian) and the M\textit{k} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.

\subsection{Combined effect of missing data parameters}
As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance are higher when all the missing data parameters have low missing data (i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have high missing data (i.e. $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-global_perparam}). However, it is important to notice that the effect of each parameter is not additive. The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$); when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}). This has important implications for building Total Evidence phylogenies (see below).

\subsection{Effects of tree inference methods}
Variation in our ability to recover the "best" tree topology depends heavily on the tree inference method (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}). However, this is the first study, to our knowledge, to compare the performance of the M\textit{k} model \citep{lewisa2001} for recovering the "best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework. Our results show that the topology of the Bayesian consensus tree is always closer to the "best" tree topology than the one of the Maximum Likelihood tree (Fig.~\ref{Fig_Results-global_perparam}). As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.

It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the "best" tree topology than Maximum Likelihood and Bayesian consensus trees. This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-permeth_perparam}). This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" trees and a set of the "best" trees involved 1000 random pairwise comparisons rather than just one. This will obviously add noise to the results for these methods.

\subsection{Practical implications}
Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections. Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "best" tree topology. We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections. Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries. However, gaps in the matrix can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}. Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "best" tree topology (Fig.~\ref{Fig_Results-paircomp_within}). Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved. One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}, so that no one individual collects all the data.

% Another practical implication of our results regards the tree building methods. Bayesian = better if chooisng one tree as backbone. However, although posterior distributions perform poorly this is more an effect of the way we compare trees (see above). Thus we do not suggest discarding posterior distribution. They are useful for comparative analyses and stuff.

@@

Using Total Evidence matrices with missing data is not a problem for recovering the "best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters. When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa. Therefore we advise filling as many gaps in Total Evidence matrices as possible. Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters as possible and to code them for as many living taxa present in the matrix as possible. Additionally, the topology of the Bayesian consensus tree, regardless the amount of missing data, was always the closer to the "best" tree topology than the Maximum likelihood trees. Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed in a Total Evidence analysis, for example when using the Tip-Dating method. % NC: Maybe one more woolly sentence.
